### 2016-12-16 Plos-one-manuscript/latex/manuscript.tex
working on figures
working

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Design}

Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
A syringe is used because the current state of FDM leaves ridges on the surface of printed parts making a airtight connection impossible. 
Micropipette designs use a piston operating principle to draw fluid into the device.
Previous 3D printed designs have gotten around this limitation by using a flexible membrane to make an airtight tube.
Then when the membrane is depressed the air is displaced allowing liquid to be drawn in.
This allows a functional pipette to be created with very few easily sourced parts.

The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer.
A 1 mL BD syringe twists to lock in the body part and is held into place by the syringe flanges.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb press. 
The pipette is spring loaded towards a set point which is adjustable by a set-screw. 
When the plunger is depressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with a lever drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to 'blow-out' the transfered fluid completely from the pipette tip.   
Additional materials required for assembly include a spring, a nut and bolt and washers.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead a combination of a barbed luer adapter and elastic tubing was used to attach the pipette tips.

Our printed pipette mimics commercial pipettes's design, function, and user operation, making it intuitive to use.
The button is pressed with the thumb until the latched position is reached.
Then the tip immersed in fluid and drawn in by actuating the lever with the palm or fingers.
Conveniently this design can also reach to the bottom of a 15 mL conical tube allowing tasks such as aspirating supernatant fluid from a cell pellet.

\subsection*{Adjusted Syringe Graduations}
Our pipette uses the air-displacement design, where a pocket of air is used to draw liquid into the pipette.
As air is an expansible fluid, this pocket of air expands due to the weight of the liquid pulling on it.
Due to this effect the graduations on the syringe are not accurate, as they are designed for measuring liquid within the syringe.
At larger volumes this effect is more pronounced resulting in the volume measured being greater then the amount of liquid pulled into the syringe.
We remedied this by creating a new scale to account for the expansion.
The scale is printed on a transparency sheet and is pasted on the syringe for accurate measurements.

\subsection*{Materials and Assembly}

\begin{table}[]
\centering
\caption{My caption}
\label{table1}
\begin{tabular}{|l|l|l|l|}
\hline
             & Unit price & Source         & Part number \\ \hline
Filiment     & \$1.63     & Makerbot       & NA          \\ \hline
1 mL Syringe & \$0.15     & BD Biosciences & 309628      \\ \hline
3 mL Syringe & \$0.73     & BD Biosciences & 309657      \\ \hline
Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
Spring       & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
Washers      & ??         & McMaster-Carr  & ??          \\ \hline
Total        & \$6.78     &                &             \\ \hline
\end{tabular}
\end{table}

% cost for each part filiment,syringe hardware
% tips used
% syringes used
% printer used and settings

%Testing taken with the original scale along with the adjusted scale is shown in table 1 and 2.

%3D parts were designed in Blender and SolidWorks and printed by Fineline Prototyping via selective laser sintering.

@@

\subsection*{Validation}
The printed pipette was validated and compared ISO 8655 and to a commercial pipette. 
The pipette was tested to see if the set point would drift between during use for example small, incremental movements of the screw or other parts causing the pipetted volume to drift.
To test the pipette the same volume was pipetted 50 times without moving or re-adjusting the set point.
A small amount of paraffin wax was applied to the screw to prevent it from slipping during each actuation. 
The printed pipette’s accuracy and precision was characterized and compared to a commercial pipette.
The printed pipette and a series of commercial pipettes were tested at volumes of 10 $\mu$L, 20 $\mu$L, 30 $\mu$L, 50 $\mu$L, 200 $\mu$L, 300 $\mu$L, 500 $\mu$L and 1000 $\mu$L. 
The printed pipette was adjusted to the the target volume by eye from the syringe graduations, and the resulting volume was measured and repeated 5 times to account for variability.
Measurements were taken from the commercial pipette in a similar way.

@@

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{ISO 8655 for 100-1000 uL comparing a commercial pipette with our printed pipette used with existing 3 mL syringe scale and an adjusted scale}
\label{table2}
\begin{tabular}{lllllll}
        &                       & Mean   & Systematic Error & \% Sys. err. & Random Error & \% Rand. err. \\
1000 uL & ISO 8655, 100-1000 uL & 1000    & 8.00             & 0.80         & 3.00         & 0.30          \\
        & Commercial Pipette    & 1002.98 & 2.98             & 0.30         & 1.72         & 0.17          \\
        & Printed Pipette       & 949.29  & -50.71           & -5.07        & 0.60         & 0.06          \\
        & Printed Pipette Scale & 1003.57 & 3.57             & 0.36         & 0.89         & 0.09          \\
        &                       &         &                  &              &              &               \\
500 uL  & ISO 8655, 100-1000 uL & 500     & 8.00             & 1.60         & 3.00         & 0.60          \\
        & Commercial Pipette    & 503.67  & 3.67             & 0.73         & 0.49         & 0.10          \\
        & Printed Pipette       & 475.99  & -24.01           & -4.80        & 4.75         & 1.00          \\
        & Printed Pipette Scale & 503.62  & 3.62             & 0.72         & 1.64         & 0.33          \\
        &                       &         &                  &              &              &               \\
200 uL  & ISO 8655, 100-1000 uL & 200     & 8.00             & 4.00         & 3.00         & 1.50          \\
        & Commercial Pipette    & 204.61  & 4.61             & 2.30         & 0.15         & 0.07          \\
        & Printed Pipette       & 186.55  & -13.45           & -6.72        & 1.31         & 0.70          \\
        & Printed Pipette Scale & 201.87  & 1.87             & 0.94         & 1.47         & 0.73          \\
        &                       &         &                  &              &              &               \\
100 uL  & ISO 8655, 100-1000 uL & 100     & 8.00             & 8.00         & 3.00         & 3.00          \\
        & Commercial Pipette    & 104.29  & 4.29             & 4.29         & 1.65         & 1.58          \\
        & Printed Pipette       & 94.02   & -5.98            & -5.98        & 4.81         & 5.12          \\
        & Printed Pipette Scale & 101.00  & 1.00             & 1.00         & 1.05         & 1.04         
\end{tabular}
%\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
%\end{flushleft}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{ISO 8655 for 30-300 uL comparing a commercial pipette with our printed pipette used with existing 3 mL syringe scale and an adjusted scale}
\label{table3}
\begin{tabular}{lllllll}
       &                       & Mean   & Systematic Error & \% Sys. err. & Random Error & \% Rand. err. \\
300 uL & ISO 8655, 30-300 uL   & 300    & 4.00             & 1.33         & 1.50         & 0.50          \\
       & Commercial Pipette    & 301.19 & 1.19             & 0.40         & 0.53         & 0.18          \\
       & Printed Pipette       & 286.91 & -13.09           & -4.36        & 0.42         & 0.15          \\
       & Printed Pipette Scale & 299.11 & -0.89            & -0.30        & 0.48         & 0.16          \\
       &                       &        &                  &              &              &               \\
200 uL & ISO 8655, 30-300 uL   & 200    & 4                & 2            & 1.5          & 0.75          \\
       & Commercial Pipette    & 200.06 & 0.06             & 0.03         & 0.46         & 0.23          \\
       & Printed Pipette       & 193.40 & -6.60            & -3.30        & 2.86         & 1.48          \\
       & Printed Pipette Scale & 200.57 & 0.57             & 0.28         & 0.86         & 0.43          \\
       &                       &        &                  &              &              &               \\
50 uL  & ISO 8655, 30-300 uL   & 50     & 4                & 8            & 1.5          & 3             \\
       & Commercial Pipette    & 49.02  & -0.98            & -1.96        & 0.10         & 0.20          \\
       & Printed Pipette       & 49.62  & -0.38            & -0.76        & 1.26         & 2.53          \\
       & Printed Pipette Scale & 48.73  & -1.27            & -2.54        & 1.11         & 2.27          \\
       &                       &        &                  &              &              &               \\
30 uL  & ISO 8655, 30-300 uL   & 30     & 4                & 13.3         & 1.5          & 5             \\
       & Commercial Pipette    & 29.08  & -0.92            & -3.06        & 0.09         & 0.31          \\
       & Printed Pipette       & 29.22  & -0.78            & -2.59        & 0.31         & 1.07          \\
       & Printed Pipette Scale & 27.78  & -2.22            & -7.41        & 1.37         & 4.93          \\
       &                       &        &                  &              &              &               \\
20 uL* & ISO 8655, 30-300 uL   & 20     & 4                & 20           & 1.5          & 7.5           \\
       & Commercial Pipette    & NA     & NA               & NA           & NA           & NA            \\
       & Printed Pipette       & 18.70  & -1.30            & -6.48        & 0.38         & 2.01          \\
       & Printed Pipette Scale & 17.94  & -2.06            & -10.29       & 1.87         & 10.42         \\
       &                       &        &                  &              &              &               \\
10 uL* & ISO 8655, 30-300 uL   & 10     & 4                & 40           & 1.5          & 15            \\
       & Commercial Pipette    & NA     & NA               & NA           & NA           & NA            \\
       & Printed Pipette       & 11.95  & 1.95             & 19.52        & 0.73         & 6.08          \\
       & Printed Pipette Scale & 7.64   & -2.36            & -23.64       & 0.38         & 4.92         
\end{tabular}
\begin{flushleft} * The 20 uL and 10 uL volumes are out of the range in this case but we wanted to demonstrate that the pipette is capable of even smaller volumes.
\end{flushleft}
\end{adjustwidth}
\end{table}

>>>>

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer.
A 1 mL or 3 mL BD syringe twists to lock in the body part and is held into place by the syringe flanges.
The 30-300 uL configuration uses a 1 mL pipette and the 100-1000 uL configuration uses a 3 mL pipette.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb press. 
The pipette is spring loaded towards a set point which is adjustable by a set-screw. 
When the plunger is depressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with a lever drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to 'blow-out' the transfered fluid completely from the pipette tip.   
Additional materials required for assembly include a spring, a nut and bolt and washers.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead a combination of a barbed luer adapter and elastic tubing is used to attach the pipette tips.
Our printed pipette mimics commercial pipettes's design, function, and user operation, making it intuitive to use.
Conveniently this design can also reach to the bottom of a 15 mL conical tube allowing tasks such as aspirating supernatant fluid from a cell pellet.

Our pipette uses the air-displacement design, where a pocket of air is used to draw liquid into the pipette.
As air is an expansible fluid, this pocket of air expands due to the weight of the liquid pulling on it.
Due to this effect the graduations on the syringe are not accurate, as they are designed for measuring liquid within the syringe.
At larger volumes this effect is more pronounced resulting in the volume measured being greater then the amount of liquid pulled into the syringe.
We remedied this by creating a new scale to account for the expansion.
The scale is printed on a transparency sheet and is pasted on the syringe for accurate measurements.

\subsection*{Assembly}

A small amount of parrifin wax is applyed to the screw to prevent slop from causing the set point to drift after each acuation.
This was effective in preventing drift.
The nut is sunk in to the hex inset in the body part.
The bolt is threaded in from the top of the body into the nut.
Springs and washers are threaded onto the plunger of the 1 mL syringe for the 30-300 uL configuration.
Springs are placed inside the 3 mL syringe for the 100-1000 mL configuration.
The plunger part is inserted in the body and the syringe assembly is pushed in and locked from the syringe flanges to the body part to complete assembly.

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 30-300 uL pipette.}
\label{table1}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	1 mL Syringe & \$0.15     & BD Biosciences & 309628      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	Washers (2)  & \$0.16     & McMaster-Carr  & 90107A012   \\ \hline
	total        & \$6.21     &                &             \\ \hline
\end{tabular}
\end{table}

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 100-1000 uL pipette.}
\label{table2}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	3 mL Syringe & \$0.73     & BD Biosciences & 309657      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	total        & \$6.63     &                &             \\ \hline
\end{tabular}
\end{table}

@@

\subsection*{Validation}
The printed pipette’s accuracy and precision was characterized and compared to a commercial pipette as well as ISO 8655.
The printed pipette was adjusted to the the target volume by eye from the syringe graduations, and the resulting volume was measured and repeated 5 times to account for transfer variability.
Data was taken for printed pipettes with existing graduations as well as with the adjusted scale.
The commerical pipette was adjusted to the desired volume and 5 transfers were taken.

@@

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{ISO 8655 for 100-1000 uL comparing a commercial pipette with our printed pipette used with existing 3 mL syringe scale and an adjusted scale}
\label{table3}
\begin{tabular}{lllllll}
        &                       & Mean   & Systematic Error & \% Sys. err. & Random Error & \% Rand. err. \\
1000 uL & ISO 8655, 100-1000 uL & 1000    & 8.00             & 0.80         & 3.00         & 0.30          \\
        & Commercial Pipette    & 1002.98 & 2.98             & 0.30         & 1.72         & 0.17          \\
        & Printed Pipette       & 949.29  & -50.71           & -5.07        & 0.60         & 0.06          \\
        & Printed Pipette Scale & 1003.57 & 3.57             & 0.36         & 0.89         & 0.09          \\
        &                       &         &                  &              &              &               \\
500 uL  & ISO 8655, 100-1000 uL & 500     & 8.00             & 1.60         & 3.00         & 0.60          \\
        & Commercial Pipette    & 503.67  & 3.67             & 0.73         & 0.49         & 0.10          \\
        & Printed Pipette       & 475.99  & -24.01           & -4.80        & 4.75         & 1.00          \\
        & Printed Pipette Scale & 503.62  & 3.62             & 0.72         & 1.64         & 0.33          \\
        &                       &         &                  &              &              &               \\
200 uL  & ISO 8655, 100-1000 uL & 200     & 8.00             & 4.00         & 3.00         & 1.50          \\
        & Commercial Pipette    & 204.61  & 4.61             & 2.30         & 0.15         & 0.07          \\
        & Printed Pipette       & 186.55  & -13.45           & -6.72        & 1.31         & 0.70          \\
        & Printed Pipette Scale & 201.87  & 1.87             & 0.94         & 1.47         & 0.73          \\
        &                       &         &                  &              &              &               \\
100 uL  & ISO 8655, 100-1000 uL & 100     & 8.00             & 8.00         & 3.00         & 3.00          \\
        & Commercial Pipette    & 104.29  & 4.29             & 4.29         & 1.65         & 1.58          \\
        & Printed Pipette       & 94.02   & -5.98            & -5.98        & 4.81         & 5.12          \\
        & Printed Pipette Scale & 101.00  & 1.00             & 1.00         & 1.05         & 1.04         
\end{tabular}
%\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
%\end{flushleft}
\end{adjustwidth}
\end{table}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{ISO 8655 for 30-300 uL comparing a commercial pipette with our printed pipette used with existing 3 mL syringe scale and an adjusted scale}
\label{table4}
\begin{tabular}{lllllll}
       &                       & Mean   & Systematic Error & \% Sys. err. & Random Error & \% Rand. err. \\
300 uL & ISO 8655, 30-300 uL   & 300    & 4.00             & 1.33         & 1.50         & 0.50          \\
       & Commercial Pipette    & 301.19 & 1.19             & 0.40         & 0.53         & 0.18          \\
       & Printed Pipette       & 286.91 & -13.09           & -4.36        & 0.42         & 0.15          \\
       & Printed Pipette Scale & 299.11 & -0.89            & -0.30        & 0.48         & 0.16          \\
       &                       &        &                  &              &              &               \\
200 uL & ISO 8655, 30-300 uL   & 200    & 4                & 2            & 1.5          & 0.75          \\
       & Commercial Pipette    & 200.06 & 0.06             & 0.03         & 0.46         & 0.23          \\
       & Printed Pipette       & 193.40 & -6.60            & -3.30        & 2.86         & 1.48          \\
       & Printed Pipette Scale & 200.57 & 0.57             & 0.28         & 0.86         & 0.43          \\
       &                       &        &                  &              &              &               \\
50 uL  & ISO 8655, 30-300 uL   & 50     & 4                & 8            & 1.5          & 3             \\
       & Commercial Pipette    & 49.02  & -0.98            & -1.96        & 0.10         & 0.20          \\
       & Printed Pipette       & 49.62  & -0.38            & -0.76        & 1.26         & 2.53          \\
       & Printed Pipette Scale & 48.73  & -1.27            & -2.54        & 1.11         & 2.27          \\
       &                       &        &                  &              &              &               \\
30 uL  & ISO 8655, 30-300 uL   & 30     & 4                & 13.3         & 1.5          & 5             \\
       & Commercial Pipette    & 29.08  & -0.92            & -3.06        & 0.09         & 0.31          \\
       & Printed Pipette       & 29.22  & -0.78            & -2.59        & 0.31         & 1.07          \\
       & Printed Pipette Scale & 27.78  & -2.22            & -7.41        & 1.37         & 4.93          \\
       &                       &        &                  &              &              &               \\
20 uL* & ISO 8655, 30-300 uL   & 20     & 4                & 20           & 1.5          & 7.5           \\
       & Commercial Pipette    & NA     & NA               & NA           & NA           & NA            \\
       & Printed Pipette       & 18.70  & -1.30            & -6.48        & 0.38         & 2.01          \\
       & Printed Pipette Scale & 17.94  & -2.06            & -10.29       & 1.87         & 10.42         \\
       &                       &        &                  &              &              &               \\
10 uL* & ISO 8655, 30-300 uL   & 10     & 4                & 40           & 1.5          & 15            \\
       & Commercial Pipette    & NA     & NA               & NA           & NA           & NA            \\
       & Printed Pipette       & 11.95  & 1.95             & 19.52        & 0.73         & 6.08          \\
       & Printed Pipette Scale & 7.64   & -2.36            & -23.64       & 0.38         & 4.92         
\end{tabular}
\begin{flushleft} * The 20 uL and 10 uL volumes are out of the range in this case but we wanted to demonstrate that the pipette is capable of even smaller volumes.
\end{flushleft}
\end{adjustwidth}
\end{table}
